>

### 2020-05-17 manuscript/manuscript.Rmd
Revise literature review and introduction
Addresses #3 and #5

Historically, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and exercise.
Over the last several years a new *social* model of health has evolved,
described by @Duhl1999 
``as an outcome of the effects of socioeconomic status, culture,
environmental conditions, housing, employment and community influences''[-@Duhl1999, p.7].
In this new paradigm, resources provided via civil infrastructure --- in
particular parks and other public green spaces --- play a crucial role in
promoting and sustaining public health [@Bedimo-Rung2005; @Wells2007; @Coutts2008].
After all, it does little to educate individuals on the importance of exercise
if their community is built in such a way that such exercise is expensive,
unenjoyable, or unsafe.

In spite of the potential for green spaces to influence public health outcomes 
and numerous studies attempting to empirically estimate the strength of these
outcomes, evidence to date has been mixed. A major challenge researchers have
faced is a proliferation of techniques for measuring access to green spaces, of
which many are plausible but lack robust theoretical bases. In this
paper, we apply a holistic and flexible measurement for park accessibility
based in activity location choice theory. This utility-based accessibility measure
compares the continuous distance to all parks in the region, weighted against
the sizes of each park and its other assorted amenities.

We apply this measurement to study the link between park accessibility and
attractiveness and census tract-level aggregate physical activity participation
and obesity rates in New York City, controlling for spatially correlated
unobserved effects, spatial spillovers of covariates, and tract-level
socioeconomic characteristics. We also demonstrate how the utility-based
accessibility metric can potentially be expanded to account for other measures
of a park's attractiveness including its observed social media activity and the
presence of various park amenities. The paper concludes with a discussion of
opportunities for future research.

# Literature Review 
\label{sec:litreview}

The large majority of extant research on the topic of park access and public health
outcomes has been focused on comparing metropolitan regions against each other.
For example, @West2012 used park data from
the Trust for Public Land’s 2010 "City Park Facts" and public health data from the
Behavioral Risk Factor Surveillance System (BRFSS) to examine the relationship
between the density of park land and physical activity and obesity rates for 67
metropolitan statistical areas in the U.S. The findings in this case conformed to 
expectation; that is, the study found a significant positive association
between park density and physical activity and a significant negative
association between park density and obesity.
@Larson2016 similarly used self-reported scores on the Gallup-Healthways
Wellbeing Index to evaluate the relationship between physical health and park
quantity, quality, and accessibility in 44 U.S. cities. While the authors found
positive associations between wellbeing and park quality and accessibility,
these relationships were not statistically significant.
Conversely, @Richardson2012 examined the relationship between total urban
greenspace and mortality rates for selected maladies; though the authors did not
find a statistically significant relationship between the quantity of urban
greenspace and mortality caused by lung cancer, diabetes, heart disease or car
accidents, they did find that all-cause mortality was --- oddly --- *higher* in greener
cities. Finally, in a meta-analysis of 20 peer reviewed journal articles exploring the
relationship between parks and objectively measured physical activity,
@Bancroft2015 found that five studies reported a significant positive
association between the two, six studies produced mixed results, and nine
studies found no association at all.

Metropolitan-level analyses such as those described above do not capture the
within-city variation in accessibility to parks that exist in many cities. Some
cities with large amounts of total greenspace may nevertheless unequally
distribute this space throughout the city, leading to areas with poor access.
Conversely, a city with a smaller overall proportion of greenspace may give all
of its citizens better access to sufficient greenspace to meet their needs or
desires for physical activity. Considering access at the neighborhood level within
a single city may also eliminate some regional or cultural fixed effects
affecting metropolitan-level analyses.

## Park Accessibility Measures

Loosely defined, the *accessibility* of an arbitrary place describes the ease
with which people can accomplish activities there. Accessibility is an abstract
concept with tempting quantifiability [@Handy1997], perhaps explaining the proliferation of
calculation mechanisms. @Dong2006 present a helpful mathematical heirarchy of
common mechanisms, which we follow here. Consider a person residing at
point $i$ in a city with parks $j \in 1 \ldots J$. An analyst might consider
point $i$ as "having access" to a park if the distance to the park $d_{ij}$ is 
within an "isochrone," or a geometric buffer defined by some threshold $D$,

@@

Variations of this isochrone-based framework are easily derived and relatively
common. 
ParkScore [@parkscore2019] is a sophisticated application of this approach where $D$ is a
carefully calculated 10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. In other applications $D$ might be more
variably defined, such as the presence or percent of green space within a
political or statistical boundary [@Mitchell2008; @Stark2014] or within a specific distance
threshold [@Kaczynski2014]. It is also possible to attenuate an isochrone with 
various factors; for instance, @Dias2019 considered road safety perception as an
extenuating factor in the effective distance to a park.

In spite of the flexibility of adapted isochrone techniques, the arbritrary
definitions necessarily imposed by researchers in its application can inhibit
holistic analysis [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of point $i$ is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_{ij})$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function and
$S_j = 1$.
@Giles-Corti2005 compare various gravity-based accessibility scores with an
isochrone specification and show that the former are more predictive of walking
behavior; that is, the attributes of a park influence walking more than
merely the distance to it. @Zhang2011 develop a population-weighted,
gravity-based accessibility to parks metric with a national scope, though they
do not examine its correlation with health outcomes.

@@

Note that the gravity formulation is itself a special case of the utility-based
specification where $\exp(X_j\boldsymbol{\beta} + \beta_d d_{ij}) = X_j f(d_{ij})$
[@Daly1982]. The distinction between gravity and utility-based specification is
meaningful, however. Primarily, it becomes possible to construct accessibility
statistics based on revealed behavioral preferences rather than calibrated or
asserted values [@Handy1997]. @Kaczynski2016 present what is effectively a 
utility-based accessibility score they call "ParkIndex", with the explicit
motivation of developing a uniformly applicable park accessibility statistic,
though they limit this statistic as a way to add heterogeneity within an
isochrone analysis.

@@

We readily acknowledge limitations in this study. As in any study conducted
with areal data, we are at risk of falling victim to the ecological inference
fallacy, where aggregate statistics mask or contradict disaggregated or
individual-level trends. A large-sample survey of individuals in New York City,
including measured physical activities and obesity would always be preferable
to the tract-level data used in this study. An ideal survey to address the
question would incorporate both sets of questions: physical activity and health
data on one hand and park use (including which parks were used and how
frequently) on the other. As no such dataset exists to our knowledge, this
tract-level aggregate analysis with asserted and transferred utility
coefficients is the possibility that remains.

@@

In both cases a common, isochrone-based analysis estimated a statistically weak
relationship with greater uncertainty. Isochrone analyses are relatively common in
the literature, perhaps because of the widespread availability of and training
in GIS software. In spite of their widespread use, isochrones are relatively
limited in terms of both their theoretical underpinnings and their flexibility
to accommodate attributes of parks beyond their proximity. And even proximity
may not be adequately handled, as the isochrone threshold may be arbitrarily
asserted by the researcher. The model we apply in this paper extends
a more comprehensive and flexible approach
for measuring the impact of park access on health outcomes. Adopting
utility-based accessibilities of the kind used in this study will allow
researchers to encompass the full range of park amenities in their accessibility
analyses, and to separate the definition of access from the study of its effects.
This will in turn enable planners to consider how multiple attributes
of a park --- from its location to its size to its amenities and beyond ---
benefit the health of the community the park serves.

>>>>

Historically, most large-scale health promotion efforts focused on
individual-level interventions intended to educate people about healthy
lifestyles and behaviors, touching on topics including diet and exercise.
Over the last several years a new *social* model of health has evolved,
described by @Duhl1999 
``as an outcome of the effects of socioeconomic status, culture,
environmental conditions, housing, employment and community influences''[-@Duhl1999, p.7].
In this new paradigm, resources provided via civil infrastructure --- in
particular parks and other public green spaces --- play a crucial role in
promoting and sustaining public health [@Bedimo-Rung2005; @Wells2007; @Coutts2008].
After all, it does little to educate individuals on the importance of exercise
if their community is designed and built in such a way that such exercise is
expensive, unenjoyable, or unsafe. It is similarly necessary for urban planners
to identify which neighborhoods have access to quality green space, and which
neighborhoods need intervention in the form of new or rehabilitated green space,
or improved transportation access to existing quality green space.

But in spite of the potential for green spaces to influence public health outcomes 
and numerous studies attempting to empirically estimate the strength of these
outcomes, evidence to date has been mixed. A major challenge researchers have
faced is a proliferation of techniques for measuring access to green spaces, of
which many are plausible but lack robust theoretical bases.

In this paper, we apply a holistic and flexible measurement for park
accessibility based in activity location choice theory. This utility-based
accessibility measure compares the continuous distance to all parks in the
region, weighted against the sizes of each park and its other assorted
amenities. We apply this measure to study the link between park accessibility
and attractiveness and census tract-level aggregate physical activity
participation and obesity rates in New York City, controlling for spatially
correlated unobserved effects, spatial spillovers of covariates, and tract-level
socioeconomic characteristics. We also demonstrate how the utility-based
accessibility metric can potentially be expanded to account for other measures
of a park's attractiveness in the form of various park amenities. The paper
concludes with a discussion of opportunities for future research.

# Literature Review 
\label{sec:litreview}
The question of how green space in the urban environment affects the physical,
emotional, and mental well-being of urban residents is not new to the academic
literature. Comprehensive reviews and analysis of this literature up to 2007
can be had from @tzoulas2007promoting, and for the subsequent decade from
@kabisch2015human. In spite of the depth of this research arena --- or perhaps
because of it --- quantitative approaches for assessing both health outcomes and
access to green space vary widely. It is therefore not surprising that in a
meta-analysis of 20 recent peer reviewed journal articles exploring the
relationship between access to parks and objectively measured physical activity,
@Bancroft2015 found that five studies reported a significant positive
association, six studies produced mixed results, and nine studies found no
association at all.

Loosely defined, the *accessibility* of an arbitrary place describes the ease
with which people can accomplish activities there.  Accessibility is an abstract
concept with tempting quantifiability [@Handy1997], but objective quantification 
may not be strictly necessary. For example, takano2002urban showed improved 
survival rates among senior citizens in Japan who self-reported having access
to qualitatively good walking spaces. Though encouraging, this result could
simply reveal self-selection: Those seniors who felt the walking space was good
were those who chose to enjoy it. Urban planners seeking to evaluate which
residents have good access to quality green space may desire objective and
therefore comparable measures.

@Dong2006 present a helpful mathematical heirarchy of
common quantitative mechanisms for calculating access, which we follow here.
Consider a person residing at
point $i$ in a city with parks $j \in 1 \ldots J$. An analyst might consider
point $i$ as "having access" to a park if the distance to the park $d_{ij}$ is 
within an "isochrone," or a geometric buffer defined by some threshold $D$,

@@

Applications of this isochrone-based framework are easily derived and relatively
common. 
ParkScore [@parkscore2019] is a sophisticated application of this approach where $D$ is a
carefully calculated 10-minute network-bound walk, paired with a demographic analysis of
areas within and outside of this threshold. @devries2003natural
showed the amount of greenspace within a 3-kilometer radius was positively associated
with a suite of self-reported health outcomes. Similarly, @devries2013streetscape
showed that streetscape greenery within 500 meters of a residence was associated with
perceived general health. Conversely, @carlson2012interactions found no relationship
between the health of senior citizens and the number of parks within 500 meters.

Another common access measure fits in this category, where distance-based
isochrones are replaced with political, geographical, or statistical boundaries.
These are especially common in comparative macroscopic studies, such as the
positive relationships found by @West2012 between density of park land in a
metropolitan area and self-reported population health indicators derived from
the Behavioral Risk Factor Surveillance System (BRFSS). On the other hand,
@Larson2016 found no relationship between physical health and park
quantity, quality, and accessibility in 44 U.S. cities and self-reported scores
on the Gallup-Healthways Wellbeing Index. At the sub-metropolitan level, 
the @Stark2014 found a significant relationship in New York City between the percent
of a residential ZIP code that is park space and objectively measured body
mass index (BMI). It is also possible to attenuate an isochrone with 
various factors; for instance, @Dias2019 considered road safety perception as an
extenuating factor in the effective distance to a park.

In spite of the flexibility of adapted isochrone techniques, the arbritrary
definitions necessarily imposed by researchers in its application can inhibit
holistic analysis [@Logan2017].
A somewhat more complete approach is the so-called "gravity" accessibility
statistic. In this case the accessibility of point $i$ is the denominator of the
gravity formulation of a trip distribution model,
\begin{equation}\label{eq:gravity}
A_i = \sum_{j = 1}^J S_j f(d_{ij})
\end{equation}
where $S_j$ is the "size" of each park literally (i.e. acres) or abstractly (i.e.
trip attractions) and $f(d_{ij})$ is a monotonically decreasing cost function,
usually a negative exponential. @Dong2006 note that the isochrone framework is a
special case of the gravity model where $f(d_{ij})$ is a binary function and
$S_j = 1$.
@Giles-Corti2005 compare various gravity-based accessibility scores with an
isochrone specification and show that the former are more predictive of walking
behavior; that is, the attributes of a park influence walking more than
merely the distance to it. @hillsdon2006relationship found no relationship between
a gravity-based accessibility term and self-reported physical activity among middle-aged individuals in Norwich, England. @Zhang2011 developed a population-weighted,
gravity-based accessibility to parks metric with a national scope, though they
do not examine its correlation with health outcomes.

@@

Note that the gravity formulation is itself a special case of the utility-based
specification where $\exp(X_j\boldsymbol{\beta} + \beta_d d_{ij}) = X_j f(d_{ij})$
[@Daly1982]. The distinction between gravity and utility-based specification is
meaningful, however. Primarily, it becomes possible to construct accessibility
statistics based on revealed behavioral preferences rather than calibrated or
asserted values [@Handy1997]. @Kinnell2006 collected a choice-based survey 
sufficient to construct a utility-based measure, though they do not do so.
Similarly, @Kaczynski2016 present what is effectively a utility-based score called
"ParkIndex", with the explicit motivation of developing a uniformly applicable
park accessibility statistic, though they limit this statistic as a way to add
heterogeneity within an isochrone analysis.

@@

We readily acknowledge limitations in this study. As in any study conducted
with areal data, we are at risk of falling victim to the ecological inference
fallacy, where aggregate statistics mask or contradict disaggregated or
individual-level trends. The socioeconomic and public health data used in this
study are only available at the tract level; projecting the tract-level data to the block
group level and computing more spatially precise accessibility logsums proved infeasible 
as the spatial lags of the independent variables introduced substantial collinearity
in the model equations. A large-sample survey of individuals in New York City,
including measured physical activities and obesity would always be preferable
to the tract-level data used in this study. An ideal survey to address the
question would incorporate both sets of questions: physical activity and health
data on one hand and park use (including which parks were used and how
frequently) on the other. As no such dataset exists to our knowledge, this
tract-level aggregate analysis with asserted and transferred utility
coefficients is the possibility that remains.

@@

In both cases a common, isochrone-based analysis estimated a statistically weak
relationship with greater uncertainty. Isochrone analyses are relatively common in
the literature, perhaps because of the widespread availability of and training
in GIS software. In spite of their widespread use, isochrones are relatively
limited in terms of both their theoretical underpinnings and their flexibility
to accommodate attributes of parks beyond their proximity. And even proximity
may not be adequately handled, as the isochrone threshold may be arbitrarily
asserted by the researcher. The model we apply in this paper extends
a more comprehensive and flexible approach
for measuring the impact of park access on health outcomes. Adopting
utility-based accessibilities of the kind used in this study will allow
researchers to encompass the full range of park amenities in their accessibility
analyses, and to separate the definition of access from the study of its effects.
This will in turn enhance the ability of planners to consider how multiple
attributes of a park --- from its location to its size to its amenities and
beyond --- benefit the health of the community the park serves.
